---
title: "analysis"
format: html
---
##Variables chosen
Questionnaire
Sleep Disorders: P_SLQ
SLD012: Sleep hours on weekdays 
Coding(3-13.5, 2 for <3, and 14 for greater than 13.5)

SLQ050: Ever told doctor you've had trouble sleeping
-----------------------------------------------
Medical conditions: P_MCQ
MCQ300a - Close relative had heart attack?

Coding: (1: Yes, 2: No, 7: Refused (1), 9: Don't know (216), .: Missing (5754))
---------------------------------------------
Food security: P_FSQ
Household food security FSDHH
Coding: 1-4: (4 is the lowest), .: Missing (1084)

Coding: (1: index is 1.3, 2: 1.3 > index <= 1.85, 3: index >1.85, 7:Refused (111), 9: don't know (482), .: Missing (1302))
-----------------------------------------
Lab data
LDL Cholesterol: P_TRIGLY
LBDLDL : LDLD cholesterol in mg/dl

Coding: (Range of values, .: Missing (440))

HDL Cholesterol P_HDL
LBDHDD: HLD in mg/dl
-----------------------------------------
Physical activity:
PAD680 - Minutes sedentary activity
Coding: Range of values, 7777: Refused (6), 9999: Don't know (60),.: Missing (17)
-----------------------------------------
Demographics:
SEQN - Respondent sequence number
RIAGENDR - Gender
Coding: 1/2/(.) Missing (0)

RIDAGEYR - Age in years at screening
0-79/80+/(.) Missing (0)
RIDSTATR - Interview/Examination status

RIDRETH3 - Race/Hispanic origin w/ NH Asian
Coding: 1/7 (7 = Other), (.) Missing (0)
-----------------------------------------
Study 2 key predictors

Close relative had a heart attack (MCQ300a) 
--------------------------------
Additional predictors
Race/Eth (RIDRETH3) + Age (RIDAGEYR) + Minutes of sedentary activity (PAD680) PAQ + Household food security (FSDHH) FSQ + Sleep hours on weekdays (SLD012) SLQ
----------------------------------
Outcome 
LDL cholesterol (LBDLDL )
##Setup
```{r, include = FALSE}
library(mosaic)
library(nhanesA)
library(tidyverse)
library(Hmisc)
library(naniar)
library(magrittr)
library(Epi)
library(janitor)
library(skimr)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(scales)
library(knitr)
library(broom)
```

```{r}
# sleep <- nhanes("P_SLQ") %>% 
#   tibble()
# 
# demo <- nhanes("P_Demo") %>% 
#   tibble()
# 
# chol <- nhanes("P_TRIGLY") %>% 
#   tibble()
# chol_hdl <- nhanes("P_HDL") %>% 
# 
# 
# saveRDS(sleep, "./data/P_SLQ.Rds")
# 
# saveRDS(demo, "./data/P_DEMO.Rds")
# 
# saveRDS(chol, "./data/P_TRIGLY.Rds")
# saveRDS(chol_hdl, "./data/P_HDL.Rds")
# 
# med_con <- nhanes("P_MCQ") %>%
#   tibble()
# 
# saveRDS(med_con, "./data/P_MCQ.Rds")
# 
# food_sec <- nhanes("P_FSQ") %>%
#   tibble()
# 
# saveRDS(food_sec, "./data/P_FSQ.Rds")
# 
# phys_acts <- nhanes("P_PAQ") %>%
#   tibble()
# 
# 
# saveRDS(phys_acts, "./data/P_PAQ.Rds")
```

```{r}
sleep_raw <- readRDS("./data/P_SLQ.Rds")

demo_raw <- readRDS("./data/P_DEMO.Rds")

chol_raw <- readRDS("./data/P_TRIGLY.Rds")

chol_hdl_raw <- readRDS("./data/P_HDL.Rds")

phys_act_raw <- readRDS("./data/P_PAQ.Rds")

food_sec_raw <- readRDS("./data/P_FSQ.Rds")

med_cond_raw <- readRDS("./data/P_MCQ.Rds")
```

```{r}
final_table <- demo_raw %>% 
  left_join(sleep_raw, by = "SEQN") %>% 
  left_join(chol_raw, by = "SEQN") %>% 
  left_join(chol_hdl_raw, by = "SEQN") %>% 
  left_join(phys_act_raw, by = "SEQN") %>% 
  left_join(food_sec_raw, by = "SEQN") %>% 
  left_join(med_cond_raw, by = "SEQN") %>% 
  select(c(SEQN, RIAGENDR, RIDAGEYR, RIDSTATR, RIDRETH3, LBDLDL, LBDHDD, SLD012, SLQ050, PAD680, MCQ300A,FSDHH)) %>% 
  mutate(RIDSTATR = as.factor(RIDSTATR)) %>% 
  mutate(RIDSTATR = recode_factor(RIDSTATR,  "Both interviewed and MEC examined" = 2, "Interviewed only" = 1)) %>% 
  mutate(SLQ050 = recode_factor(SLQ050, "Refused" = "No", r"{Don't know}" = "No")) %>% 
  filter(RIDSTATR == 2 & RIDAGEYR<80) %>% 
  mutate(age_cat = if_else(RIDAGEYR>20, 2, 1)) %>% 
  mutate(age_cat = as.factor(age_cat)) %>% 
  mutate(RIDRETH3 = as.factor(RIDRETH3))

```

## Study 1

### Analysis A

```{r, fig.width= 20}
paired <- final_table %>% 
  select(SEQN, LBDLDL , LBDHDD) %>% 
  filter(complete.cases(.))


paired %>% 
  summarise(sd_ldl = sd(LBDLDL), sd_hdl = sd(LBDHDD))%>% 
  kable()


paired_t <-  t.test(paired$LBDLDL , paired$LBDHDD, paired = TRUE, alternative = "two.sided", conf.level = 0.90, var.equal = FALSE)


tidy(paired_t) %>% 
  select(estimate, conf.low, conf.high) %>%
  kable(digits = 3)

ldl_hist <- paired %>% 
  ggplot(aes(LBDLDL))+
  geom_histogram()+ 
  labs(title = "The Distribution of LDL Cholesterol (mg/dl)")

ldl_qq <- paired %>% 
  ggplot(aes(sample = LBDLDL))+
  stat_qq() +
  stat_qq_line() +
  labs(title = "LDL QQ Plot")
  
hdl_qq <- paired %>% 
  ggplot(aes(sample = LBDHDD))+
  stat_qq() +
  stat_qq_line() +
  labs(title = "HDL QQ Plot")
  
hdl_hist <- paired %>% 
  ggplot(aes(LBDHDD))+
  geom_histogram()+ 
  labs(title = "The Distribution of HDL Cholesterol (mg/dl)")


(ldl_hist / hdl_hist) 
(ldl_qq / hdl_qq)
```

### Analysis B two means

```{r}
mean_comp <- final_table %>%
  select(RIAGENDR, LBDLDL) %>% 
  filter(complete.cases(.)) %>% 
  summarise( .by = RIAGENDR, mean_chol = mean(LBDLDL , na.rm = TRUE), n = n())

table_forPlots <- final_table %>% 
  select(SEQN, LBDLDL, RIAGENDR) %>% 
  filter(complete.cases(.))


indp_t <- t.test(mean_comp$mean_chol, mean_comp$mean_chol,conf.level = .90)

tidy(indp_t) %>% 
  select(estimate, conf.low, conf.high) %>%
  kable(digits = 3)

ldl_mean_dist <- ggplot(table_forPlots, mapping = aes(x = LBDLDL), stat = "count") +
  geom_histogram() +
  facet_grid(~ RIAGENDR) +
  labs(title = "The Distribution for LDL Cholesterol (mg/dl) by Sex")

ldl_mean_dist


```
### Analyis C

```{r}
many_groups <- final_table %>% 
  select(RIDRETH3,LBDLDL ) %>% 
  filter(complete.cases(.)) %>% 
  summarise(.by = RIDRETH3, mean_chol = mean(LBDLDL , na.rm = TRUE), n = n()) %>% 
  adorn_totals("row")
```

I don't belive ANOVA was covered in this course yet.

### Analysis D

```{r}
two_by <- final_table %>% 
  mutate(SLQ050  = fct_relevel(factor(SLQ050), "Yes", "No")) %>% 
  filter(complete.cases(RIAGENDR, SLQ050)) %$%
  table(RIAGENDR,SLQ050)

twoby2(two_by)
```

## Study 2

```{r, fig.width= 20}
var_list = c("SEQN", "LBDLDL", "RIDRETH3", "RIDAGEYR", "PAD680", "FSDHH", "SLD012", "MCQ300A")

age_breaks = seq(0,80,5)

study2_table <- final_table %>% 
  filter(RIDSTATR == 2 & between(RIDAGEYR, 21, 79)) %>% 
  filter(complete.cases(LBDLDL ,MCQ300A)) %>% 
  select(var_list) %>% 
  mutate(RIDRETH3 = as.factor(RIDRETH3),
         FSDHH = as.factor(FSDHH),
         MCQ300A = as.factor(MCQ300A)) %>% 
  mutate(RIDAGEYR = cut(RIDAGEYR, age_breaks, right = FALSE)) %>% 
  mutate(MCQ300A = recode_factor(MCQ300A, "Refused" = "No", r"{Don't know}" =  "No"))

var_check <- describe(study2_table)


ldl_dist <- study2_table %>% 
  ggplot(aes(x = LBDLDL ), stat = "count") +
  geom_histogram() 

ldl_qq <- ggplot(study2_table, aes(sample = LBDLDL )) +
  stat_qq() +
  stat_qq_line() 

ldl_dist

ldl_qq 

ldl_stats <- favstats(study2_table$LBDLDL ) %>% 
  kable()

ldl_stats

heart_check_mis <- study2_table %>% 
  filter(is.na(MCQ300A)) 

nrow(heart_check_mis)

# heart_check_pres <- study2_table %>% 
#   filter(complete.cases(MCQ300A))
# 
# heart_reth <- ggplot(heart_check_mis, mapping = aes(x = RIDRETH3), stat = "count") +
#   geom_bar() +
#   scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
#   labs(title = "Survey Respondents Missing an Answer to this Question")
# 
# heart_reth_here <- ggplot(heart_check_pres, mapping = aes(x = RIDRETH3), stat = "count") +
#   geom_bar() +
#   scale_x_discrete(guide = guide_axis(n.dodge = 2))+
#   labs(title = "Survey Respondents Who Answered this Question")
# 
# heart_age_here <- ggplot(heart_check_pres, mapping = aes(x = RIDAGEYR), stat = "count") +
#   geom_bar()+
#   scale_x_continuous(limits = c(20,80))
# 
# heart_age <- ggplot(heart_check_mis, mapping = aes(x = RIDAGEYR), stat = "count") +
#   geom_bar()
# 
# (heart_reth + heart_age) / (heart_reth_here + heart_age_here) +
#   plot_annotation(title = "Demographic Distribution of Missing vs Present Data: Relative who Experienced a Heart Attack")




ldl_check_mis <- study2_table %>% 
  filter(is.na(FSDHH)) 

ldl_check_pres <- study2_table %>% 
  filter(complete.cases(FSDHH))

ldl_reth <- ggplot(ldl_check_mis, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  labs(title = "Survey Respondents Missing an Answer to this Question")

ldl_reth_here <- ggplot(ldl_check_pres, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(title = "Survey Respondents Who Answered this Question")

ldl_age_here <- ggplot(ldl_check_pres, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

ldl_age <- ggplot(ldl_check_mis, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

(ldl_reth + ldl_age) / (ldl_reth_here + ldl_age_here) +
  plot_annotation(title = "Demographic Distribution of Missing vs Present Data: LDL Cholesterol Levels")



phys_check_mis <- study2_table %>% 
  filter(is.na(PAD680)) 

phys_check_pres <- study2_table %>% 
  filter(complete.cases(PAD680))

phys_reth <- ggplot(phys_check_mis, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  labs(title = "Survey Respondents Missing an Answer to this Question")

phys_reth_here <- ggplot(phys_check_pres, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(title = "Survey Respondents Who Answered this Question")

phys_age_here <- ggplot(phys_check_pres, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

phys_age <- ggplot(phys_check_mis, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

(phys_reth + phys_age) / (phys_reth_here + phys_age_here) +
  plot_annotation(title = "Demographic Distribution of Missing vs Present Data: Minutes of Sedentary Activity")



food_check_mis <- study2_table %>% 
  filter(is.na(FSDHH)) 

food_check_pres <- study2_table %>% 
  filter(complete.cases(FSDHH))

food_reth <- ggplot(food_check_mis, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  labs(title = "Survey Respondents Missing an Answer to this Question")

food_reth_here <- ggplot(food_check_pres, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(title = "Survey Respondents Who Answered this Question")

food_age_here <- ggplot(food_check_pres, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

food_age <- ggplot(food_check_mis, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

(food_reth + food_age) / (food_reth_here + food_age_here) +
  plot_annotation(title = "Demographic Distribution of Missing vs Present Data: Food Security Levels")




sleep_check_mis <- study2_table %>% 
  filter(is.na(SLD012)) 

sleep_check_pres <- study2_table %>% 
  filter(complete.cases(SLD012)) 
  
sleep_reth <- ggplot(sleep_check_mis, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  labs(title = "Survey Respondents Missing an Answer to this Question")

sleep_reth_here <- ggplot(sleep_check_pres, mapping = aes(x = RIDRETH3), stat = "count") +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(title = "Survey Respondents Who Answered this Question")

sleep_age_here <- ggplot(sleep_check_pres, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar()

sleep_age <- ggplot(sleep_check_mis, mapping = aes(x = RIDAGEYR), stat = "count") +
  geom_bar() 

(sleep_reth + sleep_age) / (sleep_reth_here + sleep_age_here) +
  plot_annotation(title = "Demographic Distribution of Missing vs Present Data: Hours of Weekday Sleep")

```
```{r}
set.seed(2222)
ldl_train_sample <- study2_table %>% 
  sample_frac(.7)

ldl_test_sample <- anti_join(study2_table, ldl_train_sample, by = "SEQN")




ldl_model <- lm(LBDLDL  ~ RIDRETH3 + RIDAGEYR + PAD680 + FSDHH + SLD012 + MCQ300A, data = ldl_train_sample)

ldl_model_naive <- lm(LBDLDL  ~ MCQ300A, data = ldl_train_sample)



ldl_test_performance <- augment(ldl_model, newdata = ldl_test_sample)

ldl_naive_performance <- augment(ldl_model_naive, newdata = ldl_test_sample)

tidy(ldl_model, conf.int = TRUE, conf.level = 0.90) %>%
select(term, estimate, conf.low, conf.high) %>% kable()

glance(ldl_model) %>% 
  select(nobs, r.squared, adj.r.squared,sigma, AIC, BIC) %>%
  kable(digits = 2)

glance(ldl_model_naive) %>% 
  select(nobs, r.squared, adj.r.squared,sigma, AIC, BIC) %>%
  kable(digits = 2)

favstats(~ abs(.resid), data = ldl_test_performance) %>%
  select(n, min, median, max, mean, sd) %>%
  kable(digits = 2)

favstats(~ abs(.resid), data = ldl_naive_performance) %>%
  select(n, min, median, max, mean, sd) %>%
  kable(digits = 2)

sqrt(mean(ldl_test_performance$.resid^2))

sqrt(mean(ldl_model_naive$.resid^2))


```


